\documentclass[12pt, a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{float}
\usepackage[margin=1.2in]{geometry} % margins
\usepackage{multicol}
\usepackage{wrapfig}
\usepackage[colorlinks=true, linkcolor=blue, urlcolor=cyan]{hyperref}
\usepackage{fancyhdr}
\usepackage{chemfig}
\usepackage[version=4]{mhchem}
\usepackage{chemformula}

\hypersetup {
    linkcolor=black
}

\renewcommand{\baselinestretch}{1.5}
\newcommand{\fline}{\par\noindent\rule{\textwidth}{0.1pt}}
\newcommand{\C}{\ce{C}}
\setlength{\parskip}{1em}
\addto\captionsenglish{\renewcommand{\contentsname}{Units}}

% Create commands for the molecule states
\newcommand{\gas}{_{\text{(g)}}}
\newcommand{\solid}{_{\text{(s)}}}
\newcommand{\liquid}{_{\text{(l)}}}

\pagestyle{fancy}

\fancyhead[L]{\bf\large 4U Chemistry - Notes}
\fancyhead[R]{\bf\large Semester 2}
\setlength{\headheight}{35pt}

% Create title
\title{
    \textbf{4U Chemistry:\\ Notes, Drawings, Examples}
    \author{Alexandru Stan}
    \date{Start date: February 2, 2024 \\ End date: June 27th, 2024}
}

\begin{document}

    \maketitle
    \vfill
    \newpage 

    \tableofcontents
    \fline
    \newpage

    \section{Organic Chemistry}

    \subsection{Reactions of Hydrocarbons}

    There are serveral reactions of hydrocarbons (\~20 I think?) and this is
    too low of a level to logic it out, so it will all be memorization, get
    ready.

    \subsection{Alkane Reactions}

    In general, alkanes are fairly unreactive, however, the burn very easily in combustion
    reactions, releasing a lot of energy.

    Reactants of the reaction will be the hydrocarbon and $\ce{O_{2\text{(g)}}}$ with the products
    of a complete combustion being $\ce{CO_{2\text{(g)}} + H2O_{\text{(g)}}}$

    \subsubsection{Substitution Reaction}

    During a substitution replaced a $\ce{H}$ atom with a halogen to make
    an alkyl halide. This reaction only occurs with three halogen gases $\ce{F2}$, 
    $\ce{Cl2}$, and $\ce{Br2}$. 

    This reaction can create a mixture of different isomers and can occur in multiple
    phases, making multiple substituions if needed.

    \textbf{\textit{NOTE:}} This reaction required UV light to occur

    \begin{center}
        \chemfig{C(-[:90]H)(-[:180]H)(-[:270]H) - C(-[:90]H)(-[:0]H)(-[:270]H)}
        + $\ce{Br_{2\gas} ->[\text{UV Light}]}$ \chemfig{C(-[:90]H)(-[:180]H)(-[:270]H) - C(-[:90]H)(-[:0]Br)(-[:270]H)}
        + $\ce{HBr_{\gas}}$
    \end{center}
    
    \subsection{Alcohols}

    An alcohol group is an organic compound that contains the hydroxyl $\ce{-OH}$ functional group
    the ``alcohol'' in beer and wine is really ``ethanol''

    \begin{center}
        \begin{tabular}{| c | c | c | c |}
            \hline
            $\ce{CH3OH}$ & $\ce{CH3CH2OH}$ & $\ce{CH3CH2CH2OH}$ & $\ce{CH3CH2CH2CH2OH}$\\
            \hline
            methanol & ethanol & propan-1-ol & butan-1-ol \\
            \hline
        \end{tabular}
    \end{center}

    \noindent\textbf{Some Important Alcohols:}
    \begin{itemize}
        \item Methanl - prodiced from wood, often used a solvent but is toxic
        \item Isopropanol (propan-2-yl) - Rubbing alcohol, used as an antiseptic
        \item Glycerol - Used to make fats in the body. (propan-1,2,3-ol)
    \end{itemize}

    \subsubsection{Naming Alcohols}

    When naming alcohol, the following rules need to be followed in this exact order:

    \begin{center}
        \begin{enumerate}
            \item Identify the longest chain of $\ce{C}$ that contains the $\ce{-OH}$ (hydroxyl) group
            \item Number the $\ce{C}$ atoms with the \#$1$ closest to the $\ce{-OH}$. It has prority over the alkyl 
                  groups and halogens
            \item Drop the -e ending (if two vowels are present) on alkane and add -ol. Use a number
                  if needed before the -ol
            \item Name any side branches % TODO: double check this!!
    
        \end{enumerate}
    \end{center}

    \noindent Examples:

    \begin{itemize}
        
        \item pentan-1-ol 
        
        \[
            \ce{CH3-CH2-CH2-CH2-CH2-OH}    
        \]

        \item pentan-3-ol
        
        \[
            \chemfig{CH3-CH2-CH2(-[6]OH)-CH2-CH3}    
        \]

    \end{itemize}

    \subsubsection{Aromatic Alcohols}

    The simples aromatic alcohol is a benzene ring with one hydroxyl group group
    bondede to it. Its IUPAC name is phenol

    \begin{center}
        \chemfig{*6(-=-=(-OH)-=)}
    \end{center}

    If the benzene ring has two $\ce{-OH}$ groups attached, the name is based on benzene
    and inlcudes numbers for the $\ce{-OH}$ groups

    \begin{center}
        \chemfig{*6(-=(-OH)-(-OH)=-=)}
    \end{center}

    \subsection{Primary, Secondary, and Tertiary Alchols}

    Alcohols are classified acording to where the $\ce{-OH}$ is attached. They
    are classified as primary, secondary, and tertiary alchols as stated below:

    \begin{itemize}
        \item $1^{\circ}$ (Primary Alcohol) - Hydroxyl group attached to the end 
        \item $2^{\circ}$ (Secondary Alcohol) - Hydroxyl group attached to a $\ce{C}$ that is attached
                          to two other $\ce{C}$% TODO: finish this section
        \item $3^{\circ}$ (Tertiary Alchol) - Hydroxyl group attached to a $\C$ that is attached to
                          three other $\C$
    \end{itemize}

    \subsection{Polyalcohols}

    Polyalcohols are just alcohols with more than one hydroxyl. For Nomencalture, use suffixes
    (di, tri, etc.) and numbers. The -e is kept if it followed by a consonant but is dropped
    if followed by a vowel

    \subsection{Properties of Alcohols}

    The presence of $\ce{-OH}$ group makes the molecule polar that can form hydrogen bonds. The longer
    the Carbon ($\C$) the less polar it is. Small alcohols are completely soluble in water, but the 
    solubility decreases as the length of the carbon chain increases. 

    \subsubsection{BP and MP}

    Alcohols can hydrogen bond and have higer MP and BP than 
    hydrocarbons of similar sizes

    \begin{center}
        \begin{tabular}{ | c | c | c |}
            \hline
            Molecule & Molar Mass & Boiling Point \\
            \hline
            Propane & $44$ g/mol & $-42.1^{\circ}$C \\
            \hline
            Ethanol & $46$ g/mol & $78.3^{\circ}$C \\
            \hline
        \end{tabular}
    \end{center}

    \subsection{Ethers}

    In an ether, the functional group consists of two C atoms connected to a single $\ce{O}$
    atom. The $\ce{C - O}$ bond is polar and the shape is bent, making it a polar molecule. Ethers
    are also known to be good solvents

    \begin{center}
        \chemfig{C(-[:90])(-[:180])(-[:270])-\charge{90=\:, 270=\:}{O}-C(-[:90])(-[:0])(-[:270])}\\
        \vspace*{2mm}
        An \textbf{ether}
    \end{center}

    \noindent\textbf{Some Important Ethers}

    \begin{itemize}
        
        \item Ethylene Oxide - Used in Epoxy
        \item MBTE - A gasoline aditive that helps gasoline burn better
        \item 18-crown-6 - Built by Pedersen and Cram (Nobel Prize 1987, capable of building metal ions)

    \end{itemize}

    \subsubsection{Naming Ethers}
    
    When naming ethers you need to follow a set or rules are you do alcohols and other substances, 
    although, keep in mind that ethers are made up of two parts. The rules are as follows:

    \begin{enumerate}
        \item Name the smallest alkyl group first, drop -yl and add oxy. This will be named as a side chain
        \item End the longest alkyl group last named as an alkane. This is the parent chain
    \end{enumerate}

    \begin{enumerate}
        
        \item Name the smallest alkyl group 1st, drop the -yl, and add oxy. The will
              be named as a side chain
        \item End with the longest alkyl group last named as an alkane. This is the parent chain

    \end{enumerate}

    \subsection{Alcohol and Ether Reactions}

    \subsubsection{Hydration Reaction}

    We can prepare alcohols with an addition reaction, specifically, a hydration reaction. In this
    reaction Alkenes with a double bond will accept/react with $\ce{H2O}$. One will take the $\ce{H+}$ 
    and the other the $\ce{OH-}$. 

    This reaction requires a strong acid catalyst, for which we will be using $\ce{H2SO4}$. Below is an example 
    of the reaction.

    % WRITE OUT THE propene + H2O reaction here, product will be propan-2-ol
    \begin{center}
        \schemestart
        \chemfig{\ce{H2C}(=[:330]CH(-[:30]CH3))}
        \+
        \ch{H-OH}
        \arrow(.mid east--.mid west){->[\ch{H2SO4}]}
        \chemfig{CH(-[:150]\ch{H3C})(-[6]HO)(-[:30]\ch{CH3})}
        \schemestop
    \end{center}

    
    \subsubsection{Subsitution Reaction}

    Another way an alcohol can be prepared is by susbstituting an alkyl halide with water to form an alcohol. 
    There are no reaction conditions required in this reaction, below is an example with cholocyclohexane and water

    \begin{center}
        \schemestart
        \chemfig{*6(----(-[2]Cl)--)}
        \+
        \ch{H-OH}
        \arrow(.mid east--.mid west){->}
        \chemfig{*6(----(-[2]OH)--)}
        \+
        \ch{HCl}
        \schemestop
    \end{center}

    We can also perform a substituion between an alcohol and a hydrogen halide to produce 
    an alkyl halide and water. Below is an example of 3-methylbutan-1-ol and hydrogen iodide.

    % \begin{center}
    %     \schemestart
    %     \chemfig{HC(-[:240]\ch{CH2}(-[:150]\ch{H2C}(-[2]OH)))(-[:330]\ch{CH3})(-[2]\ch{CH3})}
    %     \+
    %     \ch{H-I}
    %     \arrow(.mid east--.mid west){->} 
    %     \chemfig{CH(-[2]\ch{CH3})(-[:330]\ch{CH3})(-[:210]\ch{CH2}(-[:150]\ch{H2C}(-[2]I)))}
    %     \schemestop
    % \end{center}

    \chemfig{CH(-[2]\ch{CH3})(-[:330]\ch{CH3})(-[:210]\ch{CH2}(-[:150]\ch{H2C}(-[2]I)))}


    \subsubsection{Combustion of Alcohols}

    A combustion required $\ce{O2}$ and produces $\ce{CO2}$ and $\ce{H2O}$. With these reactions, though, 
    it is important to be careful when balancing as they can get tricky! (fractions are ok)

    \[
        \ce{C2H5OH_{\liquid} + 3O2_{\gas} -> 2CO2_{\gas} + 3H2O_{\gas}}    
    \]

    \subsubsection{Dehydration Reaction}

    In dehydration reaction results in the removal of a water form of alkene, it can be described
    as a reverse hydration reaction. Similarly to its counterpart, it also required a strong acid
    which yet again will be $\ce{H2SO4}$. An example is shown below:

    \begin{center}
        \schemestart
        \chemfig{CH(-[2]\ch{OH})(-[:330]\ch{CH3})(-[:210]\ch{H3C})}
        \+
        \arrow(.mid east--.mid west){->[\ch{H2SO4}]}
        \chemfig{CH(-[:210]\ch{H2C})(-[:330]\ch{CH3})}
        \+
        \ch{H2O}
        \schemestop
    \end{center}

    \subsubsection{Condensation Reaction}

    A condensation reaction is a way of preparing an ether from an alcohol. With this reaction, 
    the $\ce{H2O}$ is condensed out with one alcohol donating a $\ce{H+}$ and the other donating
    a $\ce{OH-}$. This reaction also required both a strong acid ($\ce{H2SO4}$) as well as extreme 
    heat. 

    \schemestart
    \chemfig{\ch{CH3CH3CH2}}
    \+
    \chemfig{\ch{CH3CH3CH2}}
    \arrow(.mid east--.mid west){->[\ch{H2SO4}]}
    \chemfig{}
    \schemestop

    \subsection{Carbonyl Group}

    Aldehydes and ketones both contain the carbonyl group. A carbonyl 
    group is a carbon atom double-bonded to an oxygen atom.

    \subsection{Aldehydes}

    In an aldehyde the carbonyl group is bonded to at least $1$ hydrogen atom (it is 
    at the end of a \ch{C} chain). The simplest two aldehydes are methanal (commonly refferd to 
    as formaldehyde) and ethanal. 

    \subsubsection{Nomenclature of Aldehydes}

    Below is how tou name aldehydes:

    \begin{enumerate}
        \item Find the longest chain with the carbonyl group
        \item Drop the -e ending and add -al
        \item When numbering the carbons on the parent chain, the 
              carbonyl group has priority
    \end{enumerate}

    \subsection{Properties of Aldehydes and Ketones}

    The carbonyl group is polar. The molecules will have dipole-dipole attraction but there will be no 
    hydrigen bonding. Their boiling points are higher than similar alkanes due to alkanes only having LDFs
    compared to the aldehydes and ketones DP-DP. Although, they have lower boiling points than similar 
    alcohols due to alcohols having hydrogen bonding. 
    
    Small aldehydes and ketones are also soluble in water,  but the solubility decreases as additional carbons 
    are added to the chain. This is because as the chain grows larger is becomes less and less polar. 

    \subsection{Reaction of Aldehydes and Ketones}

    These reactions are prepared y the controlled oxidation of an alcohol. It involves the gaining of an oxygen. 
    The reaction needs an oxidizing agent like \ch{K2Cr2)7}, \ch{H2O2}, or \ch{KMnO4}. The symbol for an 
    oxidization reaction is \ch{[O]} with the product of said reaction being and aldehyde/ketone + water.

    \begin{center}
        \schemestart
        \chemfig{C(-[:180]R)(-[:90]H)(-[:270]H)(-[:0]O(-[:0]H))}
        \arrow(.mid east--.mid west){->[\ch{[O]}]}
        \chemfig{C(-[:210]R)(=[2]O)(-[:330]H)}
        \+
        \ch{H2O}
    \schemestop
    \end{center}

    \subsubsection{$1^{\circ}$ vs $2^{\circ}$ vs $3^{\circ}$ alcohol}

    A primary alcohol in an oxidization reaction results in an aldehyde with water while a 
    secondary alcohol results in a ketone with water. Tertiary alcohols on the other hand 
    don't oxidize

    The reason that tertiary alcohols do not oxidize because for this reaction to occur
    a hydrogen need to be removed to make space for the new oxygen double bond. The tertiary
    alcohol can't do this as its bonded to other carbons and there are no hydrogen to remove.

    \subsubsection{Hydrogenation of Aledhydes and Ketones}

    This reaction occurs under conditions of high temperature and pressure, and a \ch{Pt} 
    catalyst. The reactants are hydrogen + ketone/aldehyde with its product being an 
    alcohol. Is a reverse-oxidization of alcohols where aldehydes become primary alcohols
    and ketones become secondary alcohols. 

    

\end{document}